---
title: DANL Project
subtitle: "Data-Driven Mastery: Unlocking Business Potential"
author: 
  - Jackson Boyer
  - Alban Kuka
  - Henry Socash
  - Jeremy Sun
  - Tori Pavlock

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(ggplot2)


theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```



# Introduction

This is a data.frame about the earthquakes that occurred in 2023. It provides information such as latitude, longitude, where it occurred, etc. 
```{r}
eq <- read_csv('/Users/jacksonboyer/Documents/Data Analytics/Project/jbdanl.github.io/earthquake data.csv')
```

```{r}
#| include: false
# the month with the most earthquakes occurring 
eq$Month <- rowSums(eq[,4:5])
eq$Month[1:6055] <- "NOVEMBER"
eq$Month[6056:14543] <- "OCTOBER"
eq$Month[14544:23404] <- "SEPTEMBER"
eq$Month[23405:31631] <- "AUGUST"
eq$Month[31632:40673] <- "JULY"
eq$Month[40674:48647] <- "JUNE"
eq$Month[48648:53539] <- "MAY"

densemonth <- eq %>% 
  group_by(Month) %>% 
  summarize(count =n()) %>% 
  arrange(desc(count)) %>% 
  filter(row_number() <=7)
```

# Data

The data.frame `eq` contains data about any earthquakes that have occurred during 2023. It contains data from all around the world.  


## Summary Statistics

```{r}
rmarkdown::paged_table(eq) 
```

```{r}
#| results: asis
#| echo: false
 
```


```{r}
skim(eq) %>% 
  select(-n_missing)
```

```{r}
#| include: false
summagnitude <- eq %>% 
  summarize(
    mean_magnitude = mean(Magnitude, na.rm = T), 
    sd_magnitude = sd(Magnitude),
    q1_magnitude = quantile(Magnitude, probs = .25, na.rm = T),
    median_magnitude = median(Magnitude),
    q3_magnitude = quantile(Magnitude, probs = .75, na.rm =  T),
    min_magnitude = min(Magnitude, na.rm = T),
    max_magnitude = max(Magnitude, na.rm = T)
  )
```
```{r}
skim(eq$Magnitude)
skim(eq$Depth)
```

## Top 10 countries with the most earthquakes
Below is a list of the top 10 countries that had the most earthquakes during the data collection period. There is also charts with each country corresponding to each month showing the number of earthquakes.

```{r}
#| include: false
top_10_quaked <- eq %>% 
  group_by(Country , Lands) %>% 
  summarize(count =n()) %>% 
  arrange(desc(count)) %>% 
  filter(row_number() <=10) %>% 
  head(quaked, n = 10) 
```

```{r}
rmarkdown::paged_table(top_10_quaked) 
```

```{r}
#| include: false
dense <- eq %>% 
  group_by(Month,Country,Lands) %>% 
  summarize(count =n()) %>% 
  arrange(desc(count)) %>% 
  filter(row_number() <=7) %>% 
  head(dense, n=15)
```
```{r}
ggplot(dense) +
  geom_bar(aes(y = Country, x = stat(count),
               group = 1)) +
  facet_wrap(Month ~ . , ncol = 5)
```


## Months with the most earthquakes
Below is a list of months and the corresponding number of earthquakes during that month, July being the month with the most earthquakes and May being the month with the least. The graph shows the relationship between earthquakes in each country and months earthquakes occurred in.

```{r}
rmarkdown::paged_table(densemonth) 
```
```{r}
ggplot(dense,
       aes(y = Month,
           fill = Country)) +
  geom_bar(show.legend = T)
```


## Magnitude Graphs
Below are some graphs showing the magnitude of earthquakes. 
```{r}
ggplot(eq, aes(x = log(Magnitude))) +
  geom_freqpoly(bins = 100)

ggplot(eq, aes(x = Magnitude)) +
  geom_freqpoly(bins = 100)
```


## Depth Graphs
Below are some graphs showing the depth of earthquakes. 
```{r}
ggplot(eq, aes(x = log(Depth))) +
  geom_histogram(bins = 50)

ggplot(eq, aes(x = Depth)) +
  geom_histogram(bins = 50)
```


## Magnitude Histograms
Below are histogram charts showing the magnitude of each earthquake per month.
```{r}
ggplot(eq, aes(x = Magnitude)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black", alpha = 0.7) +
  facet_wrap(~Month, scales = "free_y") +
  labs(title = "Histogram of Magnitude by Month", x = "Magnitude", y = "Count")
```


## Relationship Graph
Beloq is a graph showing the relationship between the magnitude of an earthquake and the depth it occurs at.
```{r}
ggplot(eq,
       aes(x = Magnitude, y = Depth)) +
  geom_point(alpha = .2, size = .75) +
  geom_smooth() +
  geom_smooth(method = lm, color = 'red')
```


## Earthquakes per Season
Below is a table showing the different seasons and how many earthquakes occur during that season, a graph is shown to visualize the data.
```{r}
#| include: false
eq$Season <- rowSums(eq[,4:5])
eq$Season[48648:53539] <- "SPRING"
eq$Season[23405:48647] <- "SUMMER"
eq$Season[1:23404] <- "FALL"


# which season is more earthquake prone 
denseseason <- eq %>% 
  group_by(Season) %>% 
  summarize(count =n()) %>% 
  arrange(desc(count)) %>% 
  filter(row_number() <=3)
```
```{r}
rmarkdown::paged_table(denseseason) 


ggplot(eq) +
  geom_bar(aes(x = Month, y = stat(count),
               group = 1)) +
  facet_wrap(Season ~ . , ncol = 5)
```